---
title       : Should The Space Shuttle Launch?
subtitle    : Risk of solid rocket booster o-ring failure based on ambient temperature.
author      : Justin Elszasz
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Decision: Should The Space Shuttle Launch?

* In 1986, the Space Shuttle Challenger exploded on launch, killing the crew
* Subsequent investigation showed that an o-ring seal on one of the solid rocket boosters failed
  - Low amient temperatures during launch resulted in a lack of o-ring resilience, allowing hot combustion gases to "blow-by"
* Data on 23 previous launches suggested o-rings were at risk of thermal degradation at cooler ambient temperatures
  - Data were not available for temperatures below 53 degrees Fahrenheit
  - Ambient temperature on the morning of Challenger launch was 36 degrees Fahrenheit

---

## Existing Data

* "orings" - total number of o-rings, 
* "at-risk" - number of o-rings at risk of thermal damage or blow-by
* "temp" - ambient temperature (degrees Fahrenheit)
* "leak_test_P" - pressure (pounds per square inch, or psi)
* "test_no" - test (launch) number

```{r echo=FALSE}
data <- read.table("o-ring-erosion-or-blowby.data",
                   col.names=c("orings","at_risk","temp","leak_test_P","test_no"))
head(data, 5)
```




---

## Linear Model
```{r}
training <- subset(data, !(temp > 70 & at_risk > 1)) # Remove clear outlier
model <- lm(at_risk ~ temp, data=training)
```

```{r echo=FALSE, fig.height=5, fig.width=8}
library(ggplot2)
g <- ggplot(data, aes(temp, at_risk))
g <- g + geom_vline(xintercept=36, color='blue')
g <- g + geom_abline(intercept = model$coefficients[1], slope = model$coefficients[2], size = 0.5)
g <- g + geom_point(aes(size=6, alpha=0.7))
g <- g + theme(legend.position="none")
g <- g + labs(x = "Ambient Temperature (degrees Fahrenheit) at Launch")
g <- g + labs(y = "Number of O-rings At Risk of Thermal Damage")
g <- g + annotate("text", x = 44, y = 0.5, label="Challenger Launch, 1986",color="blue")
g <- g + annotate("text", x = 75, y = 1.9, label="Assumed Outlier")
g <- g + annotate("text", x = 74, y = 0.4, label="Linear Model")
g <- g + xlim(25,85)
g <- g + ylim(0,3)
return(g)
```


---

## Application

* User selects an expected ambient temperature for space shuttle launch
* Linear regression model predicts number of o-rings at risk for thermal degradation
  - Data are an integer count of number of o-rings at risk
  - Model predicts fractional number of o-rings at risk
* Prediction used as proxy to determine whether to launch
  - If < 1, launch is safe
  - If >= 1, launch is not safe
  
### Reference
University of California - Irvine, Machine Learning Repository
[Space Shuttle O-Ring Data Set](https://archive.ics.uci.edu/ml/datasets/Challenger+USA+Space+Shuttle+O-Ring)


